### 2023-07-18 reports/report5/main.tex
Report 5 added and logbook updated

\begin{document}
\begin{titlepage}
    \begin{center}
        \LARGE
        \textbf{ELE 401 - GRADUATION PROJECT I}

@@

    \textbf{Submission Date:} 22.01.2023

@@

\begin{abstract}
    One of the most important aim of the modern technology is the sensing the environment. Living beings have sense organs to know the environment. They have brains to make fusion the sensor data and augment the environmental information taken from sense organs. For example, human standing in balance, we can say that they use the data coming from their eyes and ears together.

    People are studying to make this sensing and augmenting artificially, too. They use computers to achieve this. But computers cannot sense the environment directly. They need some devices called sensors. There are all kinds of sensors to measure different environmental quantities. For same quantity, there are different sensors using different technologies, have different precision and accuracy levels. Each sensor has its advantages and disadvantages. Even sensors that are serving a different purpose can help to augment same information other than their main. This is known as "sensor fusion"~\cite{enwiki:1115352853}. Just like humans, robots are use camera and inertial sensors to stand in balance.

    In our project, sensor fusion handles extracting depth information. Sensor fusion can be implemented using specific algorithms like EKF, CNN etc. according to quantity that will be calculated, number and types of sensors used, accuracy and performance needed. We are planning to develop a sensor fusion algorithm with the help of living organisms. As we mentioned before, living beings perform sensor fusion perfectly. We will be examining the zebra fish and how it processes data of its sense organs to produce depth information. Then we will try to develop similar algorithm or improve our algorithm with data taken from zebra fish.

    Depth information can be used for many applications like automatic driven vehicles, UAVs, Augmented Reality devices and 3D modeling of areas. Our concept in this project will be autonomous controlled SLAM Robot. Simultaneous localization and mapping (SLAM) is the computational problem of constructing or updating a map of an unknown environment while simultaneously keeping track of an agent's location within it~\cite{enwiki:1120084976}. We use a robot to move the system.

    Another concept we aim to use is cloud computing. Mobile computing devices may not match the performance requirements in some cases. To achieve more accurate depth information, complex control algorithms or neural networks will be used in our project which needs high computing power. As an alternative to the mobile computing devices, we may use cloud computing.

    From the point on we are studying on the sensor fusion. In this report we will briefly emphasize the legged robot usage for our final decisions. On the other hand, the concept ot the sensor fusion and the alternative methodologies to implement it correctly to serve the main purpose of sensor fusion, that is reliable odometry data extraction.
\end{abstract}

@@

One of the most important aim of the modern technology is the sensing the environment. Living beings have sense organs to know the environment. They have brains to make fusion the sensor data and augment the environmental information taken from sense organs. For example, human standing in balance, we can say that they use the data coming from their eyes and ears together.

People are studying to make this sensing and augmenting artificially, too. They use computers to achieve this. But computers cannot sense the environment directly. They need some devices called sensors. There are all kinds of sensors to measure different environmental quantities. For same quantity, there are different sensors using different technologies, have different precision and accuracy levels. Each sensor has its advantages and disadvantages. Even sensors that are serving a different purpose can help to augment same information other than their main. This is known as "sensor fusion"~\cite{enwiki:1115352853}. Just like humans, robots are use camera and inertial sensors to stand in balance.

In our project, sensor fusion handles extracting depth information. Sensor fusion can be implemented using specific algorithms like EKF, CNN etc. according to quantity that will be calculated, number and types of sensors used, accuracy and performance needed. We are planning to develop a sensor fusion algorithm with the help of living organisms. As we mentioned before, living beings perform sensor fusion perfectly. We will be examining the zebra fish and how it processes data of its sense organs to produce depth information. Then we will try to develop similar algorithm or improve our algorithm with data taken from zebra fish.

Depth information can be used for many applications like automatic driven vehicles, UAVs, Augmented Reality devices and 3D modeling of areas. Our concept in this project will be autonomous controlled SLAM Robot. Simultaneous localization and mapping (SLAM) is the computational problem of constructing or updating a map of an unknown environment while simultaneously keeping track of an agent's location within it~\cite{enwiki:1120084976}. We use a robot to move the system.

Another concept we aim to use is cloud computing. Mobile computing devices may not match the performance requirements in some cases. To achieve more accurate depth information, complex control algorithms or neural networks will be used in our project which needs high computing power. As an alternative to the mobile computing devices, we may use cloud computing.

@@

This project aims to construct an unmanned ground vehicle that senses the environment and decides its movements. As mentioned in the "Introduction" section, the sensor fusion and mapping algorithms take a lot of space in this project. In addition to that, there are other parts that need to be implemented like robot controls, data acquisition from sensors and cameras, decision making algorithms etc.

    \subsubsection*{A. Hardware Architecture of the Project}

    The hardware component choices are explained with consideration of advantages and disadvantages in the section 3.2 Design Constraints. Here we have our main components and specifications of them:

    \begin{enumerate}
        \item \textbf{Robot – Unitree Go1 Edu} \\
        This robot is a ready-to-go quadruped robot with built-in 5 Sets of Fish-eye Stereo Depth Cameras, Ai Post-processing, 3 Sets of Hypersonic Sensors. There are three Nvidia Jetson Nano, one Raspberry Pi 4, one STM32 Microcontroller devices inside of the Unitree Go1 Edu. These devices located in the robot to perform certain tasks. STM32 Microcontroller drives joint motors of the robot. Raspberry Pi 4 is used for both communicate with user and send commands to STM32. Two of Nvidia Jetson Nano 4GB kits and one Nvidia Jetson Nano 2 GB development kit are used for getting data from camera and sensors, processing raw data to extract dept information of stereo cameras.

        The manufacturer provides two Software Development Kits. "unitree\_legged\_sdk” for Controlling the robot, “unitree\_camera\_sdk” for getting camera data and calculating depth information.

        \item \textbf{Stereo Cameras – Unitree Camera } \\
        There are 5 sets of fish-eye stereo depth cameras located on different important locations of the robot. The known features are:

        \begin{itemize}
            \item Lens Angle $\approx$ 150x170° 
            \item HD (1280x720) , 30fps video streaming 
            \item Raw Frame 
            \item Rectangular Frame 
            \item Depth Frame 
            \item Point Cloud Image 
            \item Provided SDK 
        \end{itemize}

        \item \textbf{3D LIDAR – Velodyne 80-VLP-16-A~\cite{VelodyneLiDAR}} \\
        The VLP-16 has a range of 100m, and the sensor's low power consumption (~8W), light weight (830 grams), compact footprint (~Ø103mm x 72mm), and dual return capability make it ideal for UAVs and other mobile applications.

        Velodyne’s LiDAR Puck supports 16 channels, ~300,000 points/sec, a 360° horizontal field of view and a 30° vertical field of view, with +/- 15° up and down. The Velodyne LiDAR Puck does not have visible rotating parts, making it highly resilient in challenging environments (Rated IP67).

        Automotive Mapping UAV Security Robotics Automation

        The module provides up to 0.3 million points/second over 100Mbps ethernet cable and UDP interface. UDP packages contain distances, calibrated reflectivity information, rotation angles, synchronized time stamps (microseconds resolution).

        \item \textbf{Computation Unit – Nvidia Jetson Nano 4GB} \\
        Unitree Go1 Edu has three Nvidia Jetson Nano Development kits inside it to perform tasks like point cloud, human skeleton tracking etc. We don’t use most of the features robot has, so we decided on using one of the Nvidia Jetson Nano 4GB models in the robot with reducing its 
        workload.

        Here is some specifications of Nvidia Jetson Nano 4GB~\cite{JatsonNano}:

        \begin{itemize}
            \item GPU		:	128-core Maxwell 
            \item CPU		:	Quad-core ARM A57 @ 1.43 GHz 
            \item Memory	:	4 GB 64-bit LPDDR4 25.6 GB/s 
            \item Storage	:	microSD (not included) 
            \item Camera	:	2x MIPI CSI-2 DPHY lanes 
            \item Connectivity	:	Gigabit Ethernet, M.2 Key E 
            \item Display	:	HDMI and display port 
            \item USB		:	4x USB 3.0, USB 2.0 Micro-B 
            \item Others		:	GPIO, I2C, I2S, SPI, UART 
        \end{itemize}

        As mentioned before, one of the aims of this project is performing heavy computation on cloud. The point of implementing the algorithm on a mobile platform is to keep things simple in the first step. That provides us both simple implementation and ability to make a comparison between mobile computing and cloud computing in terms of performance, accuracy and cost.

    \end{enumerate}

    \subsubsection*{B. Software Architecture of the Project}

    Laying on the keep it simple stupid (KISS) principle, we separate our project into smaller and independent parts of development.

    \begin{enumerate}
        \item \textbf{Controlling the Robot Movement and Getting Sensor Data Using Software} \\
        There is a SDK provided by Unitree to control the robot with software called unitree\_legged\_sdk. We used the python programming language with unitree\_legged\_sdk. The details about choice of programming language are explained in the section 3.2 design constraints.

        Connection between robot and controller computer established via UDP protocol. There are two options that can provide UDP connection. One is the ethernet port of switch that connects all devices in the robot and the other is Wi-Fi access point that connected to the Raspberry Pi 4 in the robot. All devices are using a network mask 255.255.255.0, so it is important to use a static IP on the user computer.

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{UserRobotConnection.png}
            \caption{User Robot Connection: Network Mask Settings}
        \end{figure}

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{UserRPiConn.png}
            \caption{User Robot Connection: IP of Raspberry Pi}
        \end{figure}

        Unitree\_legged\_sdk allows both low-level and high-level controls. Low-level control allows user to adjust each joint manually. High-level control provides user to control robot with high-level commands like walk, turn, jump etc. We will be using high-level control for this project.

        SDK has udp class that provides us an object constructs a UDP connection with Raspberry Pi of Robot. UDP object has functions to receive, send, decode and encode udp packages.

        Another class included in the SDK is HighState. HighState objects store state information of the robot. It is updated using udp.receive(HighState) function. HighState objects includes IMU measurement information, odometer information, 2D speed information, yaw speed information, position information, body height information etc.

        To send the commands to the robot, there is a class called CMD in the SDK. CMD object is a command package that tell the robot what to do. CMD package consists of command mode, velocity, yaw speed, target position, required body height variables. Then udp.SetSend(cmd) redyes the cmd object  with encoding it to be sent. Finally the udp object sends the command to the Raspberry Pi using udp.send().

        Raspberry Pi 4 also uses unitree\_legged\_sdk to receive commands from and send state information to the user computer. It also communicates with the STM32 Microcontroller which controls all joint motors of the robot.

        \item \textbf{Getting RGB and Depth Frames from Stereo Cameras (unitree\_camera\_sdk)} \\
        As mentioned before, there are 5 sets of fish-eye stereo depth camera units to both provide an RGB frame and depth frame around the robot.

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{RobotComponentLocations.jpeg}
            \caption{Legged Robot Sensors and Other Components}
        \end{figure}

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{RobotFront.png}
            \caption{Legged Robot Head Supersensory}
        \end{figure}

        Unitree provides unitree\_camera\_sdk to access camera data of the robot. The SDK allows getting raw frames, rectangular frames, depth frames and point cloud frames.

        SDK is based on GStreamer framework and works on Nvidia Jetson Nano devices. GStreamer is a library for constructing graphs of media-handling components. Applications can take advantage of advances in codec and filter technology transparently. Developers can add new codecs and filters by writing a simple plugin with a clean, gener~\cite{gstreamer}. GStreamer creates a media pipeline that allows developers to implement their applications simple and modular.

        unitree\_camera\_sdk creates a pipeline on GStreamer and a virtual video device for each camera connected to the Jetson Nano. Developer can easily read camera information, raw frames, rectangular frames, depth frames and point cloud frames using SDK in their apps.

        \begin{figure}[H]
            \centering
            \includegraphics[width=\textwidth]{RawStereo.png}
            \caption{Raw Frame}
        \end{figure}

        \begin{figure}[H]
            \centering
            \begin{subfigure}[b]{0.4\textwidth}
                \centering
                \includegraphics[width=\textwidth]{StereoRectFrame.png}
                \caption{Rectangular Frame}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{0.4\textwidth}
                \centering
                \includegraphics[width=\textwidth]{StereoDeptFrame.png}
                \caption{Depth Frame}
            \end{subfigure}
            \caption{Worked Frames}
        \end{figure}

        Robot has some built in features like human skeleton tracking that uses cameras automatically. There are some services implemented to run these applications automatically while system is starting up. Developer should be whether kill these processes after startup or edit the services not to run processes that uses cameras.

        Dept information is calculated from stereo camera using stereo vision algorithms. First step is matching each pixel of left frame with corresponding pixel in right frame. This step includes computer vision algorithms like Sum of Squared Differences (SSD) and Sum of Absolute Differences (SAD).

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{SSDSAD.png}
            \caption{Sum of Squared Differences (SSD) and Sum of Absolute Differences (SAD)}
        \end{figure}

        After every point matched, distance of every point can be calculated using basic geometric calculations like Thales Theorem.

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{Thales.png}
            \caption{Thales Theorem}
        \end{figure}

        Stereo cameras provide three dimensional images without using moving parts and high costs. But they have disadvantages as they have advantages. Each camera has lenses to converge and diverge the light coming from different angles and make them fall onto the sensor. That means each camera has a visual angle. Other than that visual angle cannot enter the frame. As we think of the closest points of the baseline of the stereo camera that enters visual angle, it is hard to both match these points in both frames and calculate their distances from camera. Figure shows disparity with respect to distance of the point from baseline of the stereo camera.

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.6\textwidth]{DisparityOnStereo.png}
            \caption{Disparity with respect to Distance}
        \end{figure}

        \item \textbf{Developing Sensor Fusion Algorithm for 3D LIDAR and Stereo Cameras}
        \item \textbf{Simultaneously Localization and Mapping (SLAM) }
        \item \textbf{Automating SLAM with Decision Algorithm }
        \item \textbf{Moving Sensor Fusion Algorithm to Cloud}
        \item \textbf{Further Applications Like Swarm Robots with Centralized Cloud System }
    \end{enumerate}

    To accomplish this much work successfully the proof-of-concept parts should be made in a professional manner. In our earlier course work we acquired these experiences by the leading of our course supervisors. Although we will only implement software level designs, the understanding of mathematics and other engineering knowledge cannot be denied. Since we always advice with our supervisor.

    \begin{table}[H]
        \centering
        \includegraphics[width=0.65\textwidth]{GanttChart.png}
        \caption{Gantt Chart of Processes}
    \end{table}

@@

The design constraints should be identified and some discussion on how these apply to the design project and their realization are to be included in this section. You may refer to the list of some realistic design constraints that can be found in the Project Design Constraints document. Other constraints can be identified and discussed, if applicable.

@@

    We use the IEEE standards in our project. From the related pages~\cite{enwiki:1120084976} we found the good to have rules as follows:

    IEEE 829-2008, also known as the 829 Standard for Software and System Test Documentation, was an IEEE standard that specified the form of a set of documents for use in eight defined stages of software testing and system testing, each stage potentially producing its own separate type of document. The standard specified the format of these documents, but did not stipulate whether they must all be produced, nor did it include any criteria regarding adequate content for these documents. These were a matter of judgment outside the purview of the standard~\cite{enwiki:1105776890}. There will be a lot of test cases for the design. Since the system is complex, the test scenarios should be documented with known standard in order to maintenance independently from project contributors.

    A software requirements specification (SRS) is a description of a software system to be developed. It is modeled after business requirements specification (CONOPS). The software requirements specification lays out functional and non-functional requirements, and it may include a set of use cases that describe user interactions that the software must provide to the user for perfect interaction~\cite{enwiki:1117244157}. Since we have such a software system will be developed, it should lay on some standards to make the codes and all implementations reusable and modular.

    A software design description (a.k.a. software design document or SDD; just design document; also Software Design Specification) is a representation of a software design that is to be used for recording design information, addressing various design concerns, and communicating that information to the design’s stakeholders~\cite{enwiki:1105986397}. Similarly, the developed system should have standardized documentation as we said in earlier two standards. The three keywords for a good software design are: modular, reusable, and flexible.

    Of course, as the project progresses, it will be important that we comply with more detailed standards. We will continue to mention these in future reports.

@@

>>>>

\begin{document}
\begin{titlepage}
    \begin{center}
        \LARGE
        \textbf{ELE 402 - GRADUATION PROJECT II}

@@

    \textbf{Submission Date:} 18.06.2023

@@

\begin{abstract}
    Our project centers around sensor fusion, a critical aspect of artificially perceiving and enhancing the environment. By leveraging sensors, computers can acquire data on diverse environmental parameters. While each sensor has its strengths and weaknesses, even sensors designed for different purposes can contribute to augmenting the same information. Our project specifically focuses on extracting depth information through sensor fusion techniques, including EKF, CNN, and a novel iterative algorithm inspired by Papoulis Gerchberg. This algorithm operates in the frequency domain, reducing the processing workload compared to per-pixel control algorithms. The depth information obtained through our approach holds tremendous potential for a wide range of applications, including autonomous vehicles, UAVs, augmented reality devices, and 3D modeling. Additionally, we explore the integration of cloud computing to meet the computational demands of complex control algorithms or neural networks, ensuring accurate depth estimation. Our work places particular emphasis on the utilization of legged robots and highlights the significance of reliable odometry data extraction for simultaneous localization and mapping (SLAM).
\end{abstract}

@@

This report provides an overview of the completed project, which focuses on sensor fusion and SLAM operations within a fully ROS-based robot. The project aimed to develop an autonomous robotic platform capable of mapping and navigating its environment using a diverse set of sensors. The previous interim report outlined the prototype design, iterations, and evaluation results. In this final report, we will discuss the progress made since the interim report, including the additional work carried out. One significant accomplishment is the development of our sensor fusion algorithm, which has been further accelerated through the implementation of CUDA libraries. Cloud computing was also integrated into the project, and a modular Docker container structure was adopted to enhance system independence and scalability. Additionally, the SLAM process was performed using various algorithms and real-time as well as recorded datasets. Overall, this report presents a comprehensive overview of the project's accomplishments, including the development of the sensor fusion algorithm, cloud computing integration, SLAM implementation, and adherence to engineering standards and design constraints. The subsequent sections will delve into the specific details of each aspect, providing a comprehensive understanding of the completed project.

@@

Our project focuses on the development of a fully autonomous robotic platform that utilizes sensor fusion and SLAM techniques to navigate and map its environment. The primary objective is to create a robust system capable of perceiving and augmenting the surrounding environment by combining data from various sensors. By leveraging sensor fusion algorithms such as EKF, CNN, and an iterative approach inspired by Papoulis Gerchberg, we aim to extract accurate depth information.

The extracted depth information holds significant potential for a wide range of applications, including autonomous vehicles, unmanned aerial vehicles (UAVs), augmented reality devices, and 3D modeling. To handle the computational demands of complex control algorithms and neural networks for precise depth estimation, we explore the integration of cloud computing. This allows us to leverage high computing power and achieve enhanced accuracy.

Our project incorporates the use of legged robots, which offer increased mobility and adaptability in diverse environments. Additionally, reliable odometry data extraction is emphasized, as it plays a crucial role in simultaneous localization and mapping (SLAM). By ensuring accurate odometry, we can effectively construct and update a map of the unknown environment while keeping track of the robot's location within it.

Modularity, system independence, and scalability are crucial considerations for our project. To achieve these goals, we have implemented a Docker container structure that facilitates easy integration and maintenance of the system components. This approach allows for seamless expansion and upgrades as needed.

Throughout the project, we have made significant advancements in sensor fusion, algorithm development, cloud computing integration, and SLAM implementation. The combination of these elements forms a comprehensive robotic platform capable of autonomously navigating, perceiving, and mapping its surroundings. Our project not only demonstrates the practical application of sensor fusion and SLAM but also provides valuable insights and a foundation for future developments in the field.

@@

@@

Several engineering standards were followed throughout the project to ensure efficient and reliable communication, interoperability, and data transmission between the various components of the robotic system. These standards include:

\begin{itemize}
    \item TCP/IP (Transmission Control Protocol/Internet Protocol):

          The SSH (Secure Shell) protocol, based on TCP/IP, was used for secure communication between internal computers of the robot. It provides encrypted data transmission, ensuring confidentiality and integrity.

    \item UDP (User Datagram Protocol):

          UDP was utilized for transmitting Lidar data. UDP is a lightweight and connectionless protocol, well-suited for real-time applications where low latency is crucial.

    \item ROS (Robot Operating System):

          ROS was employed as the communication framework for transmitting stereo camera information and facilitating communication between different modules of the system. ROS provides a standardized messaging protocol, allowing seamless integration of various sensors and modules.

    \item Wi-Fi 802.11.ax:

          The Wi-Fi 802.11.ax standard was utilized for wireless communication. This standard offers high-speed, reliable wireless connectivity, enabling communication between the robotic platform and external devices. 
\end{itemize}

Adhering to these engineering standards ensures compatibility, efficient data transmission, and secure communication within the robotic system. By employing TCP/IP, UDP, ROS, and Wi-Fi 802.11.ax, the project meets industry best practices for communication protocols, enabling seamless interaction between components and modules of the robot.

@@

\section{PROTOTYPE}

The prototype developed for this project is a fully ROS-based robot designed to perform sensor fusion and SLAM operations. The primary purpose of building this prototype is to demonstrate the effectiveness of integrating various sensors for accurate and efficient mapping and navigation in an unknown environment. The prototype serves as a platform for testing different sensor fusion algorithms and SLAM techniques to identify the most suitable combination for real-world applications.

The prototype consists of a mobile robotic base equipped with multiple sensors, including LiDAR, camera, inertial measurement unit (IMU), and wheel encoders. These sensors work in tandem to gather data from the environment, which is then processed and combined using sensor fusion algorithms. The fused data is fed into the SLAM system, allowing the robot to construct a map of its surroundings while simultaneously keeping track of its position within that map.

Key features of the prototype include:

\begin{enumerate}
    \item ROS integration: The robot's software architecture is built on the Robot Operating System (ROS), a widely used, flexible framework for robotics development. This allows for easy integration of existing ROS packages and ensures compatibility with a wide range of sensors and algorithms.

    \item Sensor fusion: The prototype employs various sensor fusion techniques to combine the data from different sensors, resulting in a more accurate and robust representation of the environment. This helps improve the robot's overall performance in mapping and navigation tasks.

    \item SLAM implementation: The prototype integrates state-of-the-art SLAM algorithms to simultaneously localize the robot within the environment and construct a map based on the fused sensor data. This enables the robot to navigate autonomously and efficiently, even in previously unexplored areas.

    \item Modular design: The hardware and software components of the prototype are designed to be modular, allowing for easy modification, replacement, and upgrading of individual elements. This ensures that the robot can be easily adapted to accommodate new sensors or algorithms as they become available.
\end{enumerate}

LOAM (Lidar Odometry and Mapping) is a state-of-the-art algorithm for simultaneous localization and mapping (SLAM) that leverages the advantages of lidar sensors. The LOAM algorithm is designed to work in real-time and has been widely adopted in various robotic applications, including autonomous vehicles and drones, due to its robust performance and high accuracy.

The main idea behind LOAM is to separate the point cloud data generated by the lidar sensor into two components: ground points and edge points. Ground points represent the flat surfaces of the environment, while edge points correspond to the sharp features such as corners and edges. By analyzing the relative motion of these two sets of points, LOAM can estimate the robot's pose and build a detailed map of the environment.

LOAM is particularly well-suited for robotic platforms equipped with high-resolution lidar sensors, as it can handle large amounts of data efficiently and maintain accurate localization even in complex environments. Integrating LOAM into the prototype enables the robot to perform SLAM tasks with high precision, making it an ideal choice for the sensor fusion system.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\textwidth]{loam_ros_graph.png}
    \caption{Local Host LOAM ROS Node/Topic Schematic}
\end{figure}

The Robot Operating System (ROS) is a flexible and powerful framework for developing robotic applications, providing a wide range of libraries, tools, and software components for various robotic tasks. Running ROS on a Raspberry Pi (RPI) is an efficient and cost-effective way to create a compact, versatile robotic system that can take advantage of ROS's capabilities.

By utilizing an RPI as the central processing unit for the robot, the system can benefit from the RPI's low power consumption, small form factor, and extensive community support. This allows for a lightweight, portable robotic platform that can be easily extended and customized according to specific application requirements. Moreover, the RPI can run a Linux-based operating system, which is compatible with ROS, making it an ideal choice for integrating ROS into the robot's control system.

Installing and running ROS on an RPI involves setting up the necessary dependencies, configuring the ROS environment, and deploying the desired ROS packages. Once the ROS system is up and running on the RPI, the robot can utilize various ROS functionalities, such as communication between different nodes, sensor data processing, and control algorithms implementation.

Integrating ROS with the RPI enables seamless communication with other hardware components, such as sensors and actuators, through the use of ROS topics, services, and actions. This allows for efficient sensor fusion, SLAM operations, and overall system coordination. The combination of ROS and RPI offers a powerful, compact solution for developing advanced robotic applications while maintaining a low-cost and accessible platform.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\textwidth]{rpi_ros_graph.png}
    \caption{Legged Robot RPi ROS Node/Topic Schematic}
\end{figure}

\section{DESIGN PROCESS}

The design process of the prototype followed an iterative approach, involving multiple stages of development, testing, and evaluation. This methodology allowed the team to identify shortcomings in the initial design, implement modifications, and continuously improve the prototype's performance. Each iteration focused on addressing specific challenges, refining the sensor fusion techniques, and optimizing the SLAM implementation. In the following subsections, we present the steps of the design process performed in constructing the prototype, including the objectives, testing, results, and evaluation for each iteration.

\subsection{ITERATION 1: INITIAL SENSOR INTEGRATION AND ROS SETUP}

The first iteration aimed to establish a solid foundation for the project by integrating the different sensors and setting up the ROS framework for the robot. The main objectives of this iteration were:

\begin{enumerate}
    \item Selection and integration of appropriate sensors (LiDAR, camera, IMU, and wheel encoders) for the robot.
    \item Installation and configuration of ROS on the robot's embedded computer system.
    \item Setup of the communication infrastructure between the sensors, ROS nodes, and the robot's control system.
\end{enumerate}

In addition to the aforementioned advantages, the integration of the Raspberry Pi (RPI) with the Robot Operating System (ROS) allows for enhanced communication capabilities with micro-ROS systems. Micro-ROS is a lightweight version of ROS specifically designed for microcontrollers, enabling resource-constrained devices to become part of the ROS ecosystem. By incorporating micro-ROS into the system, it becomes possible to extend the functionality and modularity of the robotic platform even further.

The RPI serves as the central communication hub between the ROS nodes and the micro-ROS system. It manages the flow of data between the different subsystems and ensures seamless interaction among the various components of the robot. To facilitate this communication, the RPI is connected to a network switch, which enables efficient data transmission and distribution throughout the system.

The network switch plays a crucial role in the overall performance and reliability of the robotic system. It ensures that all components, including the RPI, micro-ROS nodes, and other connected devices, can exchange information effectively and maintain synchronized operation. By leveraging a network switch in conjunction with the RPI, the system can handle multiple data streams, balance network traffic, and reduce latency, thereby improving the overall performance of the robot.

In summary, the combination of the Raspberry Pi, ROS, micro-ROS, and a network switch provides a powerful and versatile robotic platform. The RPI acts as the central processing unit and communication hub, allowing for efficient coordination between the ROS nodes and micro-ROS systems. The use of a network switch ensures seamless data exchange among the various components of the robot, resulting in a robust, scalable, and high-performance robotic system capable of tackling a wide range of applications.

\subsubsection{TESTING AND RESULTS}

The robot's hardware and software components were assembled, and the integration of sensors with the ROS framework was tested. The communication between the sensors and ROS nodes was verified using the provided visualization tools (such as RViz) to ensure proper data flow. Preliminary tests were conducted to evaluate the robot's ability to gather data from the environment using the sensors.

\subsubsection{EVALUATION}

The results from the first iteration revealed the successful integration of sensors and the establishment of the ROS environment. However, some issues were identified, such as occasional data loss due to communication bottlenecks and the need for better synchronization between the sensors. These issues were noted and addressed in the subsequent iterations.

\subsection{ITERATION 2: SENSOR FUSION AND SLAM IMPLEMENTATION}

In the second iteration, the focus was on incorporating sensor fusion techniques to combine the data from the different sensors and implementing SLAM algorithms to enable mapping and localization capabilities. The main objectives of this iteration were:

\begin{enumerate}
    \item Development and implementation of sensor fusion algorithms to combine data from the LiDAR, camera, IMU, and wheel encoders.
    \item Selection and integration of a suitable SLAM algorithm for mapping and localization.
    \item Evaluation of the robot's performance in mapping and navigation tasks using the fused sensor data and SLAM system.
\end{enumerate}

\subsubsection{TESTING AND RESULTS}

The sensor fusion algorithms were tested to ensure the proper combination of data from the different sensors. The robot was then deployed in a controlled environment to evaluate its mapping and navigation capabilities using the SLAM system. The resulting maps were compared to the ground truth to assess the accuracy of the robot's localization and mapping.

\subsubsection{EVALUATION}

The results from the second iteration showed significant improvement in the robot's mapping and localization performance. The sensor fusion algorithms effectively combined the data from the different sensors, resulting in a more accurate and robust representation of the environment. However, some challenges remained, such as occasional drift in the robot's position estimation and the need for further optimization of the SLAM algorithm. These issues were addressed in the subsequent iterations.

\subsubsection{ITERATION 3: IMPROVING SENSOR SYNCHRONIZATION AND ROBUSTNESS}

In the third iteration, the team focused on enhancing sensor synchronization and overall system robustness to improve the robot's mapping and localization capabilities. The main objectives of this iteration were:

\begin{enumerate}
    \item Refine sensor synchronization to minimize data loss and inconsistencies between sensor readings.
    \item Implement error handling and recovery mechanisms to increase the system's robustness.
    \item Optimize the SLAM algorithm to reduce drift and improve localization accuracy.
\end{enumerate}

\subsubsection{TESTING AND RESULTS}

The improved sensor synchronization was tested by analyzing the consistency of sensor data and verifying the data alignment using visualization tools. The robot was then deployed in a controlled environment, including areas with challenging conditions, such as uneven terrain, dynamic obstacles, and poor lighting. The robot's performance in mapping, localization, and navigation tasks was assessed, taking into consideration the implemented error handling and recovery mechanisms.

\subsubsection{EVALUATION}

The results of the third iteration showed a notable improvement in sensor synchronization and overall system robustness. The robot was able to handle challenging environmental conditions more effectively and recover from errors more efficiently. However, some areas for further optimization were identified, such as refining the SLAM algorithm for better performance in dynamic environments and enhancing the robot's ability to cope with sensor noise.

\subsection{ITERATION 4: DYNAMIC ENVIRONMENT HANDLING AND NOISE REDUCTION}

The fourth iteration aimed to address the remaining challenges by improving the robot's performance in dynamic environments and reducing the impact of sensor noise. The main objectives of this iteration were:

\begin{enumerate}
    \item Adapt the SLAM algorithm to better handle dynamic environments and moving obstacles.
    \item Implement noise filtering techniques to reduce the impact of sensor noise on the robot's mapping and localization performance.
    \item Evaluate the overall performance of the robot, including its robustness, adaptability, and accuracy in various environmental conditions.
\end{enumerate}

\subsubsection{TESTING AND RESULTS}

The modified SLAM algorithm and noise filtering techniques were tested in multiple environments, including static and dynamic settings. The robot was deployed in a controlled environment with moving obstacles and varying lighting conditions to assess its ability to adapt and maintain accurate mapping and localization performance.

\subsubsection{EVALUATION}

The results from the fourth iteration demonstrated the robot's improved performance in dynamic environments and its enhanced ability to handle sensor noise. The robot was able to maintain accurate mapping and localization despite the presence of moving obstacles and varying environmental conditions. The noise filtering techniques effectively reduced the impact of sensor noise on the robot's performance. With these improvements, the prototype was considered ready for further real-world testing and potential future applications.

\section{FINAL DESIGN}

    \subsection{MEETING THE CONSTRAINTS AND ENGINEERING STANDARDS}

    The final design of our project successfully met the design constraints outlined in the project description. We ensured compliance with the specified economy, environment, manufacturability, and sustainability constraints.

        \subsubsection{Economy}

        Throughout the project, we maintained strict control over costs to meet the project's economic constraints. By leveraging the Unitree Go1 Edu robot's internal Nvidia Jetson Nano 4GB devices for computing power, we were able to reduce expenses compared to using external computing units. Additionally, the use of readily available sensors and components helped keep the overall costs within budget.

        \subsubsection{Environment}

        The environmental impact of our project was considered at various stages of development. We focused on using battery-powered systems, minimizing power consumption, and adhering to legal regulations regarding radiation and noise pollution. By selecting energy-efficient components and optimizing algorithms, we aimed to reduce the ecological footprint of the robotic platform.

        \subsubsection{Manufacturability}

        The prototype design prioritized manufacturability by utilizing readily available products and components. This approach minimized the need for high-level manufacturing technologies and complex assembly processes. The modular design of the robot and its compatibility with existing ROS packages also facilitated the integration of different hardware and software components.

        \subsubsection{Sustainability}

        Our project aimed to establish a sustainable foundation for future developments. By emphasizing the mathematical foundations, software abstractions, and comprehensive engineering solutions, we laid the groundwork for future advancements and upgrades. The modular design of the robot, adherence to engineering standards, and consideration of safety measures ensured the long-term sustainability and reliability of the robotic platform.

        The project also met the engineering standards and best practices to ensure the reliability, maintainability, and safety of the robotic platform. We followed software engineering principles, such as modular design, code documentation, version control, and code reviews. The adoption of the ROS framework provided a standardized and modular approach to software development, enhancing interoperability and code reusability. Safety measures were implemented to prevent potential hazards, and documentation and reporting were maintained throughout the project to ensure traceability and knowledge dissemination.

    \subsection{COST ANALYSIS}

    The cost analysis of our project includes both initial costs and potential future costs for production and maintenance. The initial costs include the purchase of components and materials required for building the prototype, while the future costs involve scaling up production and maintaining the robotic platform.

        \subsubsection{Initial Costs}

        The initial costs for our project primarily consisted of the Unitree Go1 Edu robot and additional sensors. The Unitree Go1 Edu robot, including its embedded computing units and sensors, had a cost of approximately \$ 18,000. The cost breakdown of the additional sensors used in the project is as follows:

        \begin{itemize}
            \item LiDAR: Velodyne 80-VLP-16-A 3D LiDAR - \$ 2500.
            \item Camera: StereoLabs ZED Stereo Camera - \$ 1000.
        \end{itemize}

        The total initial cost of the prototype, including the robot and additional sensors, amounted to approximately \$ 22,500.

        \subsubsection{Future Costs}

        Future costs for production and maintenance may vary depending on the scale of deployment and specific application requirements. Some potential future costs to consider include:

        \begin{itemize}
            \item Production Costs: Scaling up production would involve the cost of manufacturing additional robotic platforms, including the necessary components and assembly processes. These costs would depend on the production volume and any customization required for specific applications.
            \item Maintenance Costs: Maintenance costs would include periodic hardware inspections, software updates, and component replacements. The costs would depend on the frequency and complexity of maintenance tasks, as well as any warranty or support agreements.
            \item Upgrades and Expansion: As technology advances and new sensors or algorithms become available, there may be additional costs associated with upgrading or expanding the capabilities of the robotic platform. These costs would depend on the specific upgrades and their integration complexity.
        \end{itemize}

        It is important to note that the costs mentioned above are estimates based on current market prices and may vary depending on factors such as location, supplier, and future market conditions.

        \subsubsection{Cost-Benefit Analysis}
